---
title: "ACS Data Download & Messy Mapping"
author: "Susan E. Chen, Austin Burcham, Ryan Jacobs, Tim Pierce"
Last Update: "7/21/2021"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
    theme: readable
colorlinks: true
urlcolor: blue
linkcolor: blue
citecolor: blue
anchorcolor: blue
toccolor: blue
fontsize: 12pt
---

```{r setup, message = FALSE }
knitr::opts_chunk$set(echo = TRUE)
#Packages to load & manipulate ACS data.
# If you don't have the following packages installed yet, uncomment and run the lines below
#install.packages("tigris")
#install.packages("dplyr")
#install.packages("leaflet")

library(tigris)
library(dplyr)
library(leaflet)
library(tidycensus)
library(tidyverse)
library(dplyr)
library(stringr)
library(tigris)
```
# Overview of extracting ACS data
# Set your working directory 
```{r}
setwd("/Users/tspierce/Desktop/DSPG/ALMV/Data")

```

# Get a Census API key
To use the tidycensus package in `R` you will need a Census API key. You can get an API key at https://api.census.gov/data/key_signup.html. You should save this API key in the file `.Renviron` in your home directory. You can do this using `census_api_key=KEYGIVENBYCENSUS`. To confirm you have read in your census key you can type  `Sys.getenv("census_api_key")` after you have read in the API key.

```{r set_up_API, }
#readRenviron("~/.Renviron") #readin Census API key
```

# Setting up the list of counties to pull

We need to match FIPS state codes to STATES and then merge them with the subset of counties we want to pull from the ACS. The following code reads in the list of all FIPS codes for state and county that I want to pull from the ACS. I made this list in excel and read in the CSV below.
```{r counties_to_pull}
#Create a nice list of FIPS state and county. Use `fip_list` below to pull variables and to map.
counties<-read.csv("All_Mining.csv", header=T)
#Note Yellowstone National Park did not have ACS data so drop it.
counties <- filter(counties,FIPS!= 30113)

fip_list<-sprintf("%05d",counties$FIPS)
fips  <- data.frame(fip_list)
state_list<-unique(sprintf("%02d", counties$state_code))
states <- data.frame(state_list)
```

# Getting ACS data
The code chunk creates a function to pull one variable from the American Community Survey (ACS). The function pulls the variable specified, in the example below the variable is B19301.  The function is `almv_acs_table()` and to call it you use almv_acs_table("B19301"). It pulls the estimate (which I rename to something more user friendly) and also the geometry need to map the variable.

I also use the fip_list to pull variables from the ACS dataset as I want a subset of counties in the US. Use the variable table number from the ACS website (@Ryan-replace http link to website here please https://data.census.gov/cedsci/table?q=S2801&g=0500000US01007&tid=ACSST5Y2019.S2801&hidePreview=true) e.g. `B19301`. 

Note: make sure that geometry=TRUE option is set below because you need this to make maps. Here I only pull this one time and save in the PerCapita dataframe.

This function works if there is only one variable in the online table: 

```{r pull_ACS_data}
almv_acs_table <- function(varcode){
  get_acs(geography="county",
          state=state_list, cache_table = TRUE, geometry = TRUE,
          table = varcode, 
          year=2019, survey = "acs5") %>%
    filter(GEOID %in% fip_list)  %>%    select(GEOID, NAME, estimate, geometry)
  
}

#Call the function to pull data from ACS
PerCapita <- almv_acs_table("B19301") %>%  rename(PerCapInc=estimate)
#write.csv(PerCapita$GEOID,'PerCapita.csv')

```

For some variables it will pull many rows of variables. You can specify the row and column to pull with an underscore for column and then an underscore for row. For example column 3 and row 1 would be `S1810_C03_001`. Not I set `geometry=F` because I already have the geometry I need.

```{r pull_ACS_data_var}
almv_acs_var <- function(varcode){
  get_acs(geography="county",
          state=state_list,
          variables =varcode,
          year=2019, 
          survey = "acs5",
          cache_table = TRUE,
          geometry = F) %>%
    filter(GEOID %in% fip_list) %>%
    select(NAME, estimate)
}
#keep only variable of interest at the county level
disability <- almv_acs_var("S1810_C03_001") %>%  rename(Pct.Dis=estimate)
HI_insured <-  almv_acs_var("S2701_C03_001") %>%  rename(Pct.HI=estimate)
Unemployment <- almv_acs_var("S2301_C04_001") %>%  rename(Pct.Unemp=estimate)
Housing  <- almv_acs_var("S2502_C04_001") %>%  rename(Housing.OwnOcc=estimate)
Housing2  <- almv_acs_var("S2502_C01_001") %>%  rename(Housing.Total=estimate)

#Calculate your own estimate
SCounty.Pop <- almv_acs_var("B15003_001")  %>%  rename(SCountyPop2=estimate)
Ed0 <- almv_acs_var("B15003_025")  %>%  rename(Ed.PhD=estimate)
Ed1 <- almv_acs_var("B15003_024")  %>%  rename(Ed.Prof=estimate)
Ed2 <- almv_acs_var("B15003_023")  %>%  rename(Ed.Mast=estimate)
Ed3 <- almv_acs_var("B15003_022")  %>%  rename(Ed.Bach=estimate)
Ed4 <- almv_acs_var("B15003_021")  %>%  rename(Ed.Assoc=estimate)
Ed5 <- almv_acs_var("B15003_020")  %>%  rename(Ed.SColl=estimate)
Ed6 <- almv_acs_var("B15003_019")  %>%  rename(Ed.SCollLT1=estimate)
Ed7 <- almv_acs_var("B15003_018")  %>%  rename(Ed.GED=estimate)
Ed8 <- almv_acs_var("B15003_017")  %>%  rename(Ed.HSDip=estimate)
Ed9 <- almv_acs_var("B15003_016")  %>%  rename(Ed.HSNoDip=estimate)
#less than high school??? - create this category of less than hs by subtractnig countypop2 - all other cats.

CountyHH <- almv_acs_var("S2801_C01_001") %>%  rename(County.HH=estimate)
HHCompdev <- almv_acs_var("S2801_C02_002") %>%  rename(HH.Pct.compdev=estimate)
HHComputer <- almv_acs_var("S2801_C02_003") %>%  rename(HH.Pct.computer=estimate)
HHInternet <- almv_acs_var("S2801_C02_012") %>%  rename(HH.Pct.internet=estimate)
HH_BB <- almv_acs_var("S2801_C02_014") %>%  rename(HH.Pct.BroadBand=estimate)
HH_NoInternet <- almv_acs_var("S2801_C02_019") %>%  rename(HH.Pct.Nointernet=estimate)


CountyMedianAge <- almv_acs_var("S0101_C01_032") %>%  rename(County.MedAge=estimate) 
CountyMedianInc <- almv_acs_var("S1901_C01_013") %>%  rename(County.MedInc=estimate) 


Age1  <- almv_acs_var("S0101_C02_002") %>%  rename(Pct.Under5=estimate) 
Age2  <- almv_acs_var("S0101_C02_003") %>%  rename(Pct.Bet5_9=estimate) 
Age3  <- almv_acs_var("S0101_C02_004") %>%  rename(Pct.Bet10_14=estimate) 
Age4  <- almv_acs_var("S0101_C02_005") %>%  rename(Pct.Bet15_19=estimate) 
Age5  <- almv_acs_var("S0101_C02_006") %>%  rename(Pct.Bet20_24=estimate) 
Age6  <- almv_acs_var("S0101_C02_007") %>%  rename(Pct.Bet25_29=estimate) 
Age7  <- almv_acs_var("S0101_C02_008") %>%  rename(Pct.Bet30_34=estimate) 
Age8  <- almv_acs_var("S0101_C02_009") %>%  rename(Pct.Bet35_39=estimate) 
Age9  <- almv_acs_var("S0101_C02_010") %>%  rename(Pct.Bet40_44=estimate) 
Age10  <- almv_acs_var("S0101_C02_011") %>%  rename(Pct.Bet45_49=estimate) 
Age11  <- almv_acs_var("S0101_C02_012") %>%  rename(Pct.Bet50_54=estimate) 
Age12  <- almv_acs_var("S0101_C02_013") %>%  rename(Pct.Bet55_59=estimate) 
Age13  <- almv_acs_var("S0101_C02_014") %>%  rename(Pct.Bet60_64=estimate) 
Age14  <- almv_acs_var("S0101_C02_015") %>%  rename(Pct.Bet65_69=estimate) 
#Calculate Age.70plus

LT20K_NoInternet<- almv_acs_var("S2801_C02_023") %>%  rename(LT20K.Pct.Nointernet=estimate)
LT2075K_NoInternet<- almv_acs_var("S2801_C02_027") %>%  rename(LT2075K.Pct.Nointernet=estimate)
GT75K_NoInternet<- almv_acs_var("S2801_C02_031") %>%  rename(GT75K.Pct.Nointernet=estimate)

```
# Merging Your Data Horizontally

Merge variables pulled above horizontally by county and put into one dataset using base R. Note the join variable is `NAME` because it is the only unique variable in all the datasets that we are merging. List the datasets you want to join in `list()` as you see below. 
```{r merge_all}
all_ACS5 <- Reduce(function(x, y) merge(x, y, all=TRUE), list(PerCapita,disability,HI_insured,Unemployment,Housing,Housing2,Age1,Age2,Age3,Age4,Age5,Age6,Age7,Age8,Age9,Age10,Age11,Age12,Age13,Age14,Ed0,Ed1,Ed2,Ed3,Ed4,Ed5,Ed6,Ed7,Ed8,Ed9,SCounty.Pop,CountyHH,HHCompdev,HHComputer,HHInternet,HH_BB,HH_NoInternet,     LT20K_NoInternet,LT2075K_NoInternet,GT75K_NoInternet,CountyMedianAge,CountyMedianInc))
```

# Merging in other datasets
Merge in the Coal_Region variable so we can filter on this variable if we only want to get one region.
The chunk below create a new dataframe with the `Coal_Region` attached.

Note: GeoID is the FIPS Code for county and can be used for the merge

```{r}
coalcounties  <- data.frame(Coal_Region=counties$coal_region,GEOID=sprintf("%05d", counties$FIPS))

all_ACS5 <- Reduce(function(x, y) merge(x, y, all=TRUE), list(coalcounties,all_ACS5))
```
The function below will use the dataframe all_ACS5 to draw a map for one variable. Note: make sure to specify the variable with the dataset in front i.e. all_ACS5$PerCapita.  

# Cleaning up the desktop
```{r cleanup}
ls()
#remove all objects that contain "Age"
#rm(list = ls()[grepl("Age", ls())])
#rm(list = ls()[grepl("Ed", ls())])
#rm(list = ls()[grepl("Hous", ls())])
#rm(list = ls()[grepl("HH", ls())])
#rm(list = ls()[grepl("fips", ls())])
#rm(list = ls()[grepl("NoInt", ls())])
#rm("HI_insured", "disability")
```

# Draw a map with data
Here is a simple way to draw a map. Better ways exist but this is a rough guide. 

## State Borders
To put an overlay of the STATE on top of your counties you need to save the STATE geometry for making maps. You will use this in your maps below. This will change depending on your subsets of states. 

You can get state borders using the states() function in the `TIGRIS` package

Right now my state borders are messed up because they are different for each region. This needs to be fixed by the user.

All maps need to be cleaned up to have a color blind friendly palate with nice breaks.

## One Map
```{r mapAvar}
state_borders <- states(cb = T)
state_borders <- state_borders %>% 
                  filter(STATEFP %in% state_list)

zap <- ggplot(all_ACS5$geometry) + geom_sf(aes(fill=all_ACS5$Pct.Dis, color = all_ACS5$Pct.Dis)) + scale_color_viridis_c() + scale_fill_viridis_c()+ 
    coord_sf(datum = NA) +
    theme_minimal()
zap
```

# Create a function to map 
The makes mapping easier. These maps are UGLY - please make them better
```{r}
#specify function to create maps easily
ACS_map <- function(mapvar,mapdata,mapborders){
    estimate=mapvar
    borders=mapborders
    ggplot(mapdata) +
    geom_sf(data = borders, color = 'black', fill = 'grey')+ geom_sf(aes(fill=estimate, color = estimate)) + scale_color_viridis_c() + scale_fill_viridis_c()+ 
    coord_sf(datum = NA) +
    theme_minimal()
  }
#The function below will map the variable
DMAP1 <- ACS_map(all_ACS5$Pct.Dis,all_ACS5$geometry,state_borders)
DMAP1 #show the map

DMAP2 <- ACS_map(all_ACS5$PerCapInc,all_ACS5$geometry,state_borders)
DMAP2 #show the map

```

# Saving and Exporting the R dataframe
At this point I would like to use this dataframe in leaflet to produce maps, etc.  

```{r}
save(all_ACS5,file="all_ACS5.Rda")

```
In another R script you can load the data that you saved above
```{r load_data}
load("all_ACS5.Rda")
```

A new way to do this is to save the dataframe as an *.rds file. You can then read it in and save it with another name
```{r}
#saveRDS(all_ACS5, "all_ACS5.rds")
#mydat <- readRDS("all_ACS5.rds")
```
You can also save multiple objects to a file and then read them in using the function save() i.e. can be used to save one or more R objects to a specified file (in .RData or .rda file formats). The function can be read back from the file using the function load().
```{r}
# Save multiple objects & use this one to map
save(state-List, all_ACS5, file = "ACS_Objects.RData")
# To load the data again
#load("junk.RData")
```

#Save your entire workspace
It’s a good idea to save your workspace image when your work sessions are long. This can be done at any time using the function `save.image()`.

That stores your workspace to a file named .RData by default. This will ensure you don’t lose all your work in the event of system reboot, for instance.
```{r}
save.image(file = "pull_ACS5.RData") 
```
To then load the workspace use: 
```{r}
load("pull_ACS5.RData")
```


# Colorblind Friendly packages
Below I the viridis packages to make a color blind friendly palate. The options for this package are" "viridis", "magma", "inferno", or "plasma".
```{r}
# Create a color palette with handmade bins.
library(viridis)

```



# Get the spatial data (tigris) for the leaflet map
```{r}

#Get the list of states you want counties for 
nstate_list<-as.numeric(state_list)
# grab the spatial data (tigris)
county_poly <- counties(state=nstate_list,cb=TRUE)

#Do the merge (tigris)
poly_all_ACS5<- geo_join(county_poly, all_ACS5, "GEOID", "GEOID")

poly_all_ACS5 <- na.omit(poly_all_ACS5)   

```

# Creating a map with leaflet
See: http://zevross.com/blog/2015/10/14/manipulating-and-mapping-us-census-data-in-r-using-the-acs-tigris-and-leaflet-packages-3/#set-up-the-packages
1. Specify bins for the legend & a colorblind friendly palette.

```{r}
#specify the bin breaks
mybins <- c(0,10000,20000,30000,40000,50000,60000,70000)
#specify the default color
mypalette <- colorBin(palette="inferno", domain=all_ACS5$PerCapInc, na.color="transparent", bins=mybins)

leaflet(data = poly_all_ACS5) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~mypalette(poly_all_ACS5$`PerCapInc`),
    stroke=TRUE,
    weight = 1,
    smoothFactor = 0.2,
    opacity = 1.0,
    fillOpacity = 0.7, 
    label=paste("County: ",poly_all_ACS5$GEOID, ", Value: ",poly_all_ACS5$PerCapInc),
    highlightOptions = highlightOptions(color = "white",
                                        weight = 2,
                                        bringToFront = TRUE)) %>%
  addLegend(pal=mypalette, position = "bottomright",
            values = ~poly_all_ACS5$PerCapInc,
            opacity = 0.5, title = "Per Capita Income") %>%
  addPolylines(data = county_poly, color = "black", 
               opacity = 0.2, weight = 1)

```

2. Or a different way to draw the map with a different baselayer.
```{r}

pal <- colorNumeric(
  palette = "inferno",
  domain = poly_all_ACS5$PerCapInc
)
map4<-leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = poly_all_ACS5, 
              fillColor = ~pal(PerCapInc), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2) %>%
  addLegend(pal = pal, 
            values = poly_all_ACS5$PerCapInc, 
            position = "bottomright", 
            title = "Percent of Households") 

map4
```


